--- 
title: "Aides Drug"
author: "Sandra Iovino!"
date: "June 2024"

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)

druggies = select(triUmp, c(treatment, infected, DrugUser))
druggies = filter(druggies, DrugUser == "yes")

``` 

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

$H_0$: The treatment is not affected to the infected drug users 
$H_A$: the treatment is effected to the infected drug User 

# Methods

Treatment is a categorical variable   
Infected is a categorical variable  
I used the helper file   
CAT ~ CAT   

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->


```{r}
barchartGC(~infected + treatment,data=druggies)
barchartGC(~infected + treatment,data=druggies, type="percent")
```

 When running this barchart it shows that drug users who are in placebo are more likely to get aids then those who have the Aid's treatment.    
 
### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment,data=druggies)
rowPerc(table1)
colPerc(table1)
```
This numerical results shows people who had AID and where in the actual drug treatment were more effective than those who were in the placebo.     

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi square shows P-Value of .0000001 which is less than 0.05, so we reject the null hypothesis that the treatment is not affected to the infected drug users.  

When looking into the Fisher's Exact Test it shows P-Value .00000004 which is also less than 0.05, and we also reject the null hypothesis that the treatment does not work to infected drug users. 

## Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Our question was trying to see if the new trial medication was highly effective or people who are postive in AIDS or is placebo was better. When doing the calculations we noticed that infected drug users responded better to the new study treat. This means that it would be less likely for them to have aids then those who are on the placebo treatment.  